---
title: '#3 Life Factor'
author: "Lu Liang"
date: "2025-02-27"
output: html_document
---

3. Lifestyle Factors and Sleep:
How does caffeine intake affect sleep start time or sleep quality?
Does screen time before bed impact total sleep hours or sleep quality?
Is there a link between exercise duration and sleep quality or total sleep hours?


## Load the Libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(shiny)
library(shinydashboard)
require(janitor)
library(palmerpenguins)
```

```{r}
sleep <- read_csv("~/Desktop/BIS15W2025_group7/Lu's Folder (#3)/sleep_cycle_productivity.csv") %>% 
  clean_names()
sleep$sleep_quality <- as_factor(sleep$sleep_quality)
```

```{r}
summary(sleep)
```
```{r}
sleep_data <- sleep %>%
  mutate(Age_Group = cut(age, 
                         breaks = c(18, 22, 30, 45, 60, 100), 
                         labels = c("18-22", "23-30","31-45", "46-60", "60+"),
                         right = FALSE))  # Ensure age bins are non-overlapping


```


##1How does caffeine intake affect sleep start time or sleep quality?
_no effect_
```{r}
sleep_data %>% 
  select(sleep_start_time,caffeine_intake_mg,Age_Group) %>% 
  ggplot(aes(x=caffeine_intake_mg,y=sleep_start_time))+
  geom_point(alpha=0.8,color="skyblue")+
  geom_smooth(method = "lm", color = "red")+
  facet_grid(Age_Group~.)
```
```{r}
sleep %>% 
  select(sleep_start_time,caffeine_intake_mg,age,gender) %>% 
  filter(age=="18", gender=="Female") %>% 
  ggplot(aes(x=caffeine_intake_mg,y=sleep_start_time))+
  geom_point(alpha=0.8,color="skyblue")+
  geom_smooth(method = "lm",se=T, color = "red")
```

```{r}
sleep$sleep_quality <- as_factor(sleep$sleep_quality)
sleep %>% 
  select(caffeine_intake_mg,sleep_quality) %>% 
  ggplot(aes(x=sleep_quality,y=caffeine_intake_mg,fill=sleep_quality))+
  geom_boxplot()
```
```{r}
sleep %>% 
  select(caffeine_intake_mg,sleep_quality) %>%
  group_by(sleep_quality) %>% 
  summarise(mean_intake=mean(caffeine_intake_mg)) %>% 
  arrange(-mean_intake)
```

```{r}
sleep %>% 
  select(caffeine_intake_mg,sleep_quality) %>%
  group_by(sleep_quality) %>% 
  summarise(mean_intake=mean(caffeine_intake_mg)) %>% 
  arrange(-mean_intake) %>% 
  ggplot(aes(x=reorder(sleep_quality,mean_intake),y=mean_intake))+
  geom_col()+
  coord_flip()

```
##2Does screen time before bed impact total sleep hours or sleep quality?
_no effect_
```{r}
sleep %>% 
  select(sleep_start_time,screen_time_before_bed_mins,gender,age) %>%
  filter(age=="50") %>% 
  ggplot(aes(x=screen_time_before_bed_mins,y=sleep_start_time))+
  geom_point(alpha=0.8,color="skyblue")+
  geom_smooth(method = "lm", color = "red")
```

```{r}
sleep %>% 
  select(screen_time_before_bed_mins,sleep_quality) %>% 
  ggplot(aes(x=sleep_quality,y=screen_time_before_bed_mins,fill=sleep_quality))+
  geom_boxplot()
```

```{r}
sleep %>% 
  select(screen_time_before_bed_mins,sleep_quality) %>%
  group_by(sleep_quality) %>% 
  summarise(mean_screen=mean(screen_time_before_bed_mins)) %>% 
  arrange(-mean_screen)
```

```{r}
sleep %>% 
  select(screen_time_before_bed_mins,sleep_quality) %>%
  group_by(sleep_quality) %>% 
  summarise(mean_screen=mean(screen_time_before_bed_mins)) %>% 
  arrange(-mean_screen) %>% 
  ggplot(aes(x=reorder(sleep_quality,mean_screen),y=mean_screen))+
  geom_col()+
  coord_flip()
```

```{r}
sleep %>% 
  select(exercise_mins_day,sleep_quality) %>% 
  ggplot(aes(x=sleep_quality,y=exercise_mins_day))+
  geom_boxplot()
```


```{r}
sleep %>% 
  select(screen_time_before_bed_mins,sleep_quality) %>%
  group_by(sleep_quality) %>% 
  summarise(mean_screen=mean(screen_time_before_bed_mins)) %>% 
  arrange(-mean_screen) %>% 
  ggplot(aes(x=sleep_quality,y=mean_screen))+
  geom_point()
```
```{r}
sleep_data %>% 
  group_by(Age_Group) %>% 
  summarise(mean_intake=mean(caffeine_intake_mg),
            mean_sleep=mean(sleep_start_time),
            mean_total=mean(total_sleep_hours))
  
```


```{r}
sleep %>% 
  group_by(age) %>% 
  summarise(mean_intake=mean(caffeine_intake_mg),
            mean_sleep=mean(sleep_start_time),
            mean_total=mean(total_sleep_hours))
  
```



```{r}
summary(sleep$sleep_start_time)
```
```{r}
cor(sleep$caffeine_intake_mg, sleep$sleep_start_time)
cor(sleep$caffeine_intake_mg, sleep$sleep)
```

```{r}
sleep_caffeine_group <- sleep %>%
  mutate(caffeine_bin = cut(caffeine_intake_mg,
                            breaks = c(-1,0, 50, 100, 150, 200, 250, 300),
                            labels = c("0","0-50" ,"51-100","101-150","151-200","201-250","251-300"))) 

```


```{r}
sleep_caffeine_group %>% 
  group_by(caffeine_bin) %>% 
  summarise(mean=mean(sleep_start_time))
```


```{r}
sleep_caffeine_group %>% 
  group_by(caffeine_bin) %>% 
  summarise(mean = mean(as.numeric(sleep_start_time)))  %>% 
  mutate(caffeine_bin = as.numeric(caffeine_bin)) %>% 
  ggplot(aes(x=caffeine_bin,y=mean))+
  geom_point()+
  geom_smooth(method="lm")
```

```{r}
sleep_caffeine_group <- sleep_caffeine_group %>%
  mutate(caffeine_bin = factor(caffeine_bin, levels = c("0", "0-50", "51-100", "101-150", "151-200", "201-250", "251-300")))

sleep_caffeine_summary <- sleep_caffeine_group %>% 
  group_by(caffeine_bin) %>% 
  summarise(mean_sleep_start = mean(sleep_start_time, na.rm = TRUE))  

ggplot(sleep_caffeine_summary, aes(x = caffeine_bin, y = mean_sleep_start)) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "lm") + 
  labs(x = "Caffeine Intake Level", y = "Mean Sleep Start Time") +
  theme_minimal()
```


