---
title: "test"
author: "Lu Liang"
date: "2025-03-06"
output: html_document
---
# Load libraries
```{r}
library(ggplot2)
library(dplyr)
```
# Read data (replace with your file path)
```{r}
data <- read.csv("sleep_cycle_productivity.csv")
```

# 1. Caffeine Intake vs. Sleep Start Time (Scatter plot with regression line)
```{r}
ggplot(data, aes(x = Caffeine.Intake..mg., y = Sleep.Start.Time)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Caffeine Intake vs. Sleep Start Time",
       x = "Caffeine Intake (mg)",
       y = "Sleep Start Time (24h decimal)") +
  theme_minimal()
```

# Spearman correlation



# 2. Caffeine Intake vs. Sleep Quality (Boxplot with binned caffeine)
```{r}
data <- data %>%
  mutate(Caffeine_Bin = cut(Caffeine.Intake..mg.,
                            breaks = c(0, 50, 150, 300),
                            labels = c("0-50 mg", "51-150 mg", "151-300 mg")))

ggplot(data, aes(x = Caffeine_Bin, y = Sleep.Quality)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Caffeine Intake vs. Sleep Quality",
       x = "Caffeine Intake",
       y = "Sleep Quality (1-10)") +
  theme_minimal()

```

# 3. Screen Time vs. Total Sleep Hours (Scatter plot)
```{r}
ggplot(data, aes(x = Screen.Time.Before.Bed..mins., y = Total.Sleep.Hours)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "darkgreen", se = FALSE) +
  labs(title = "Screen Time Before Bed vs. Total Sleep Hours",
       x = "Screen Time Before Bed (mins)",
       y = "Total Sleep Hours") +
  theme_minimal()
```


# 4. Screen Time vs. Sleep Quality (Boxplot)
```{r}
data <- data %>%
  mutate(Screen_Time_Bin = cut(Screen.Time.Before.Bed..mins.,
                               breaks = c(0, 30, 60, 200),
                               labels = c("0-30 mins", "31-60 mins", "61+ mins")))

ggplot(data, aes(x = Screen_Time_Bin, y = Sleep.Quality)) +
  geom_boxplot(fill = "salmon") +
  labs(title = "Screen Time Before Bed vs. Sleep Quality",
       x = "Screen Time Before Bed",
       y = "Sleep Quality (1-10)") +
  theme_minimal()

```

# 5. Exercise Duration vs. Sleep Quality (Scatter plot)
```{r}
ggplot(data, aes(x = Exercise..mins.day., y = Sleep.Quality)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "purple", se = FALSE) +
  labs(title = "Exercise Duration vs. Sleep Quality",
       x = "Exercise Duration (mins/day)",
       y = "Sleep Quality (1-10)") +
  theme_minimal()

```

# Spearman correlation for key relationships
```{r}
cor_exercise_sleep_quality <- cor.test(data$Exercise..mins.day., data$Sleep.Quality, method = "spearman")
cor_screen_sleep_quality <- cor.test(data$Screen.Time.Before.Bed..mins., data$Sleep.Quality, method = "spearman")

cat("Key Correlations:\n")
cat("Exercise vs. Sleep Quality (ρ):", round(cor_exercise_sleep_quality$estimate, 2), "\n")
cat("Screen Time vs. Sleep Quality (ρ):", round(cor_screen_sleep_quality$estimate, 2))
```

