---
title: "Group Project"
output: html_document
date: "2025-02-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(shiny)
library(shinydashboard)
library(shinythemes)
```


```{r}
sleeping <- read.csv("/Users/monobuoy/Desktop/BIS15W2025_group7/sleep_cycle_productivity.csv") %>% clean_names()
sleeping
```
1. Sleep and Productivity:
Does total sleep duration (total_sleep_hours) influence productivity scores (productivity_score) and stress at around 7 hours of total sleep
```{r}
 plot_productivity <- sleeping %>% 
  filter(gender != "Other") %>% 
  filter(total_sleep_hours >= 6.75 & total_sleep_hours <= 7.25) %>% 
  ggplot(aes(x = productivity_score, fill = gender)) +
  geom_density(alpha=0.5) +
  labs(title = "Productivity Score Density",
       x = "Productivity Score", 
       y = "Density")
ggsave("productivity_score_density.png", plot = plot_productivity, width = 8, height = 5, dpi = 300)

plot_stress <- sleeping %>% 
  filter(gender != "Other") %>% 
  filter(total_sleep_hours >= 6.75 & total_sleep_hours <= 7.25) %>% 
  ggplot(aes(x = stress_level, fill = gender)) +
  geom_density(alpha=0.5) +
  labs(title = "Productivity Score Density",
       x = "Productivity Score", 
       y = "Stress")
ggsave("stress_level_density.png", plot = plot_stress, width = 8, height = 5, dpi = 300)
```
Does total sleep duration (total_sleep_hours) influence productivity scores (productivity_score) and stress at around 9.5 hours of total sleep
```{r}
 plot_productivity2 <- sleeping %>% 
  filter(gender != "Other") %>% 
  filter(total_sleep_hours >= 9.25 & total_sleep_hours <= 9.75) %>% 
  ggplot(aes(x = productivity_score, fill = gender)) +
  geom_density(alpha=0.5) +
  labs(title = "Productivity Score Density",
       x = "Productivity Score", 
       y = "Density")
ggsave("productivity_score_density.png", plot = plot_productivity2, width = 8, height = 5, dpi = 300)

plot_stress2 <- sleeping %>% 
  filter(gender != "Other") %>% 
  filter(total_sleep_hours >= 9.25 & total_sleep_hours <= 9.75) %>% 
  ggplot(aes(x = stress_level, fill = gender)) +
  geom_density(alpha=0.5) +
  labs(title = "Productivity Score Density",
       x = "Productivity Score", 
       y = "Stress")
ggsave("stress_level_density.png", plot = plot_stress2, width = 8, height = 5, dpi = 300)
```

#caffeine intake vs productivity
```{r}
library(shiny)
library(dplyr)
library(ggplot2)

# Define UI
ui <- fluidPage(
  
  titlePanel(tags$h1("Sleep, Productivity & Well-being", style = "font-size: 36px;")),
  
  sidebarLayout(
    sidebarPanel(
      sliderInput("sleep_filter", 
                  "Filter by Total Sleep Hours:", 
                  min = 4, max = 10, value = 6, step = 0.5)
    ),
    
    mainPanel(
      fluidRow(
        column(6, plotOutput("plot_productivity", width = "100%", height = "700px")),  # Left density plot
        column(6, plotOutput("plot_stress", width = "100%", height = "700px"))  # Right density plot
      )
    )
  )
)

# Define Server
server <- function(input, output, session) {
  
  session$onSessionEnded(stopApp) # Stop the app when closed


#Productivity Score by Gender
  output$plot_productivity <- renderPlot({
    sleeping %>% 
      filter(gender != "Other") %>% 
      filter(total_sleep_hours >= as.numeric(input$sleep_filter) - 0.25 & 
             total_sleep_hours <= as.numeric(input$sleep_filter) + 0.25) %>%  # Noise range ±0.25
      ggplot(aes(x = productivity_score, fill = gender)) +
      geom_density(alpha = 0.5) + 
      scale_fill_manual(values = c("#e74c3c", "#3498db")) +
      labs(title = paste("Productivity Score Density (Sleep:", input$sleep_filter, "±0.25 hrs)"), 
           x = "Productivity Score", 
           y = "Density") +
      theme_minimal()
  })
  
#Stress Level by Gender
  output$plot_stress <- renderPlot({
    sleeping %>% 
      filter(gender != "Other") %>% 
      filter(total_sleep_hours >= as.numeric(input$sleep_filter) - 0.25 & 
             total_sleep_hours <= as.numeric(input$sleep_filter) + 0.25) %>%  # Noise range ±0.25 
      
      ggplot(aes(x = stress_level, fill = gender)) +
      geom_density(alpha = 0.5) +  
      scale_fill_manual(values = c("#2ecc71", "#9b59b6")) +
      labs(title = "Stress Level Density by Gender", 
           x = "Stress Level", 
           y = "Density") +
      theme_minimal()
  })
}

# Run the Shiny App
shinyApp(ui, server)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r setup, include=FALSE}
library(shiny)
library(ggplot2)
library(shinydashboard)
library(reshape2)
library(rlang)
library(RColorBrewer)
library(shinythemes)
library(shinycssloaders)
```

```{r setup, include=FALSE}
# UI
ui <- dashboardPage(
  dashboardHeader(title = HTML("<b>Sleep</b>")),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Scatter Plot", tabName = "scatter", icon = icon("braille")),
      menuItem("Boxplot", tabName = "boxplot", icon = icon("box")),
      menuItem("Density Plot", tabName = "density", icon = icon("chart-area")),
      menuItem("Histogram", tabName = "histogram", icon = icon("chart-bar")),
      menuItem("Column Plot", tabName = "columnplot", icon = icon("bar-chart"))
    ),
    selectInput("xvar", "Select X-axis Variable (Numeric | Categorical):", 
                choices = list(
                  "Numeric Variables" = c("Caffeine.Intake..mg.", "Screen.Time.Before.Bed..mins.", "Total.Sleep.Hours", "Exercise..mins.day."),
                  "Categorical Variables" = c("Age", "Mood.Score", "Stress.Level", "Sleep.Quality", "Productivity.Score")
                )),
    selectInput("yvar", "Select Y-axis Variable (Numeric | Categorical):", 
                choices = list(
                  "Numeric Variables" = c("Total.Sleep.Hours", "Caffeine.Intake..mg.", "Screen.Time.Before.Bed..mins.", "Exercise..mins.day."),
                  "Categorical Variables" = c("Age", "Mood.Score", "Stress.Level", "Sleep.Quality", "Productivity.Score")
                )),
    sliderInput("ageRange", "Select Age Range:", min = 18, max = 80, value = c(18, 80)),
    selectInput("gender", "Select Gender:", choices = c("All", "Male", "Female", "Other"), selected = "All")
  ),
  dashboardBody(
    tags$head(
      tags$style(HTML("body { background-color: #B7C3D0; }
        .shiny-notification {
          font-size: 18px;
          font-weight: bold;
          color: white;
          background-color: #FF6B6B;
          border: 2px solid #D32F2F;
          padding: 15px;
          text-align: center;
        }"))
    ),
    tabItems(
      tabItem(tabName = "scatter", fluidRow(box(withSpinner(plotOutput("scatterPlot")), width = 12))),
      tabItem(tabName = "boxplot", fluidRow(box(withSpinner(plotOutput("boxPlot")), width = 12))),
      tabItem(tabName = "density", fluidRow(box(withSpinner(plotOutput("densityPlot")), width = 12))),
      tabItem(tabName = "histogram", fluidRow(box(withSpinner(plotOutput("histogramPlot")), width = 12))),
      tabItem(tabName = "columnplot", fluidRow(box(withSpinner(plotOutput("columnPlot")), width = 12)))
    )
  )
)

# Server
server <- function(input, output) {
  
  df <- read.csv("sleep_cycle_productivity.csv", check.names = TRUE)
  
  filtered_df <- reactive({
    data <- df[df$Age >= input$ageRange[1] & df$Age <= input$ageRange[2], ]
    if (input$gender != "All") {
      data <- data[data$Gender == input$gender, ]
    }
    return(data)
  })
  
  output$scatterPlot <- renderPlot({
    if (input$xvar %in% c("Age", "Mood.Score", "Stress.Level", "Sleep.Quality", "Productivity.Score")) {
      showNotification("⚠️ Warning: The selected X-axis variable is categorical and may not be suitable for a scatter plot!", type = "error", duration = 5)
    }
    if (input$yvar %in% c("Age", "Mood.Score", "Stress.Level", "Sleep.Quality", "Productivity.Score")) {
      showNotification("⚠️ Warning: The selected Y-axis variable is categorical and may not be suitable for a scatter plot!", type = "error", duration = 5)
    }
    plot_title <- paste('Scatter Plot:', input$xvar, 'vs', input$yvar)
    ggplot(filtered_df(), aes(x = !!sym(input$xvar), y = !!sym(input$yvar))) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "lm", color = "blue", se = FALSE) +
      theme_minimal() +
      labs(title = plot_title, x = input$xvar, y = input$yvar)
  })
  
  output$boxPlot <- renderPlot({
    if (input$xvar %in% c("Caffeine.Intake..mg.", "Screen.Time.Before.Bed..mins.", "Total.Sleep.Hours", "Exercise..mins.day.")) {
      showNotification("⚠️ Warning: The selected X-axis variable is numeric and may not be suitable for a boxplot!", type = "error", duration = 5)
    }
    plot_title <- paste('Boxplot of', input$xvar, 'vs Total Sleep Hours')
    ggplot(filtered_df(), aes(x = as.factor(!!sym(input$xvar)), y = !!sym("Total.Sleep.Hours"))) +
      geom_boxplot(fill = "lightblue", color = "darkblue") +
      theme_minimal() +
      labs(title = plot_title, x = input$xvar, y = 'Total Sleep Hours')
  })
  
  output$densityPlot <- renderPlot({
    if (!is.null(input$yvar)) {
      showNotification("⚠️ Warning: The density plot does not require a Y-axis variable!", type = "error", duration = 5)
    }
    plot_title <- paste('Density Plot of', input$xvar)
    ggplot(filtered_df(), aes(x = !!sym(input$xvar), fill = ..density..)) +
      geom_density(alpha = 0.6, color = "darkblue", fill = "lightblue") +
      theme_minimal() +
      labs(title = plot_title, x = input$xvar, y = 'Density')
  })
  
  output$histogramPlot <- renderPlot({
    if (input$xvar %in% c("Caffeine.Intake..mg.", "Screen.Time.Before.Bed..mins.", "Total.Sleep.Hours", "Exercise..mins.day.")) {
      showNotification("⚠️ Warning: The selected X-axis variable is numeric and may not be suitable for a histogram!", type = "error", duration = 5)
    }
    plot_title <- paste('Histogram of', input$xvar)
    ggplot(filtered_df(), aes(x = as.factor(!!sym(input$xvar)))) +
      geom_bar(fill = "lightblue") +
      theme_minimal() +
      labs(title = plot_title, x = input$xvar, y = 'Count')
  })
  
  output$columnPlot <- renderPlot({
    if (input$xvar %in% c("Caffeine.Intake..mg.", "Screen.Time.Before.Bed..mins.", "Total.Sleep.Hours", "Exercise..mins.day.")) {
      showNotification("⚠️ Warning: The selected X-axis variable is numeric and may not be suitable for a column plot!", type = "error", duration = 5)
    }
    plot_title <- paste('Column Plot of', input$xvar, 'vs', input$yvar)
    ggplot(filtered_df(), aes(x = as.factor(!!sym(input$xvar)), y = !!sym(input$yvar), fill = as.factor(!!sym(input$xvar)))) +
      geom_col() +
      theme_minimal() +
      labs(title = plot_title, x = input$xvar, y = input$yvar)
  })
}

# Run the app
shinyApp(ui, server)

```
```

